---
title: "Big Mac Index Analysis"
author: "Pragya Verma"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE, warning = FALSE, message = FALSE)

library(dplyr)     # For dataframe manipulation
library(simcf)     # For panel functions and simulators
library(tseries)   # For ADF unit root test
library(plm)       # Econometrics package for linear panel models
library(nlme)      # Package for estimation - RE models
library(plm)       # Package for estimation - FE models
library(kableExtra)# Beautifying the table
library(ggplot2)   # package for plotting
library(gridExtra)
library(grid)
library(MASS)      # for random nv
```

```{r}
df <- read.csv('cleanedData/processed_data_I.csv')
attach(df)
head(df)
```

Plotting the time series, 
```{r, echo = FALSE}
# Plotting the time series
ggplot(data = df, aes(x=year, y=big_mac_index)) + 
  geom_line(aes(colour=iso_a3))
```
From above plot we can see that there is no trend present in the data. 

```{r}
countrylist <- unique(Country.Code)
new.country <- countrylist[-20]
```

So we will check the stationarity for each time series country wise.
```{r, echo=FALSE}
n = 20
# Check for a unit root in each country
PPtest.pvalues <- rep(0,n)
adftest.pvalues <- rep(0,n)
for (i in 1:length(new.country)) {
    currcty <- new.country[i]

    # Check PP unit root test, omitting errors due to short series
    curPP <- try(PP.test(big_mac_index[new.country==currcty ])$p.value)
    if (any(class(curPP)=="try-error")) curPP <- NA
    PPtest.pvalues[i] <- curPP
    
    # Check ADF unit root test
    curadf <- try(adf.test(big_mac_index[new.country==currcty ])$p.value)
    if (any(class(curadf)=="try-error")) curadf <- NA
    adftest.pvalues[i] <- curadf
  }

```

Plotting the histograms of the p-values - 
```{r, echo=FALSE}
hist(PPtest.pvalues)          # Plot a histogram of the p-values

hist(adftest.pvalues)         # Plot a histogram of the p-values
```


Checking for stationarity,

Performing Im-Pesaran-Shin panel unit root tests for Big Mac index assuming fixed intercepts-
```{r, echo = FALSE}
ts <- with(df, data.frame(split(big_mac_index, as.character(year)))) 

purtest(ts, pmax = 4, exo = "intercept", test = "ips")
```
```{r, echo = FALSE}
ts <- with(df, data.frame(split(big_mac_index, as.character(year)))) 

purtest(ts, pmax = 4, exo = "trend", test = "ips")
```

We can see that the data is stationary in nature.



## Plots 

```{r}
# group Big mac index by country
country.bigmac <- aggregate(df$big_mac_index, list(df$iso_a3), FUN=mean)

# Barplot
ggplot(country.bigmac, aes(x=Group.1, y=x)) + 
  geom_bar(stat = "identity") +
  xlab("Country") + ylab("Big Mac Index") + ggtitle("Dollar Valuation by Country")
```


```{r}
# group Big mac index by country
year.bigmac <- aggregate(df$big_mac_index, list(df$year), FUN=mean)

# Barplot
ggplot(year.bigmac, aes(x=Group.1, y=x)) + 
  geom_line(stat = "identity") +
  xlab("Country") + ylab("Big Mac Index") + ggtitle("Annual Dollar Valuation")
```


```{r}
# basic scatterplot
gdp <- ggplot(df, aes(x=GDP.per.capita, y=big_mac_index)) + 
    geom_point()

ca <- ggplot(df, aes(x=Curr.Acc.Balance, y=big_mac_index)) + 
    geom_point()

labcomp <- ggplot(df, aes(x=labor.comp.per.capita, y=big_mac_index)) + 
    geom_point()

ggb <- ggplot(df, aes(x=GGB, y=big_mac_index)) + 
    geom_point()

grid.arrange(grobs = list(gdp, ca, labcomp, ggb), ncol=2, 
             main="Absolute valuation vs Big Mac determinants graph")
```
```{r, echo = FALSE}
ts <- with(df, data.frame(split(relative_PPP, as.character(year)))) 

purtest(ts, pmax = 4, exo = "trend", test = "ips")
```

```{r}
adf.test(df$relative_PPP)
```

```{r}
pp.test(df$relative_PPP)
```


## ACF and PACF plots 

```{r}
# Look at the ACF for each country
for (i in 1:length(new.country)) {
    currcty <- new.country[i]
    acf(df$big_mac_index[new.country==currcty])
}
```

```{r}
# Look at the PACF
for (i in 1:length(new.country)) {
    currcty <- new.country[i]
    pacf(df$big_mac_index[new.country==currcty])
}
```

## Model 1

### Pooled OLS
```{r}
model <- big_mac_index ~ GDP.per.capita + Curr.Acc.Balance + GGB + labor.comp.per.capita

pooled_ols_plm <- plm(model, data = df[df$iso_a3 != "USA" & df$iso_a3 != "EUK", ], 
                      index = c("Country.Code", "year"), 
                      effect = "individual", model = "pooling")

summary(pooled_ols_plm)
```

### Fixed Effects
```{r}
fe_model_plm <- plm(model, data = df[(df$iso_a3 != "USA") & (df$iso_a3 != "EUK"), ], 
                    index = c("Country.Code", "year"), 
                    effect = "individual", model = "within")

summary(fe_model_plm)
```



### Random effects
```{r}
re_model_plm <- plm(model, data = df[(df$iso_a3 != "USA")  & (df$iso_a3 != "EUK"), ], 
                    index = c("Country.Code", "year"), 
                    effect = "individual", model = "random")

summary(re_model_plm)
```


```{r}
#hausmen test
phtest(fe_model_plm, re_model_plm)
```
The Hausmen test tells us that the appropriate model is the fixed effect model.

```{r}
# averaging by year
summarize.df <- df%>%group_by(year)%>%summarise_all("mean")
summarize.df
```

## Model 2

```{r}
df$big.mac.afforability <- (df$GDP.per.capita/df$dollar_price) / 365
head(df)
```

### Pooled OLS
```{r}
model.2 <- big_mac_index ~ big.mac.afforability + Curr.Acc.Balance + GGB + labor.comp.per.capita

pooled_ols_plm <- plm(model.2, data = df[df$iso_a3 != "USA" & df$iso_a3 != "EUK", ], 
                      index = c("Country.Code", "year"), 
                      effect = "individual", model = "pooling")

summary(pooled_ols_plm)
```


### Fixed Effects
```{r}
fe_model_plm <- plm(model.2, data = df[(df$iso_a3 != "USA") & (df$iso_a3 != "EUK"), ], 
                    index = c("Country.Code", "year"), 
                    effect = "individual", model = "within")

summary(fe_model_plm)
```

## Random Effect
```{r}
re_model_plm <- plm(model.2, data = df[(df$iso_a3 != "USA") & (df$iso_a3 != "EUK"), ], 
                    index = c("Country.Code", "year"), 
                    effect = "individual", model = "random")

summary(re_model_plm)
```

















